# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, include = FALSE}
library(ggplot2)
library(tidyverse)
library(scales)
library(vcd)
library(plotly)
library(readr)
library(GGally)
library(parcoords)
```

```{r}
film_clean <- read_csv("data_exports/5-3-film-clean-for-analysis.csv")


film_clean <- film_clean %>%
  mutate(award_mcu = factor(award_mcu, levels = c("Non-MCU", "MCU", "Academy Award Nominee", "Academy Award Winner")),
         decade = cut(as.numeric(year), breaks = seq(from = 1960, to = 2030, by = 10), 
                        include.lowest = TRUE, right = FALSE, labels = FALSE),
         decade = factor(decade, levels = c(1, 2, 3, 4, 5, 6, 7),
                         labels = c("60's", "70's", "80's", "90's", "00's", "10's", "20's")))

```

```{r}

mcu_theme <- function(){
  font <- "Helvetica"
  theme_minimal() +
  theme(title = element_text(family = font, size = 20),
        axis.title = element_text(family = font, size = 18),
        axis.text.x = element_text(family = font, size = 16),
        axis.text.y = element_text(family = font, size = 16),
        legend.title = element_text(family = font, size = 18),
        legend.text = element_text(family = font, size = 16)
        )
}

colors <- c("#000000", #"Black"
           "#233f92", #"Dark Cornflower Blue"
           "#44a0e3", # "Carolina Blue"
           "#C4D6FA", # Light Blue
           "#8b6b87", # "Chinese Violet"
           "#d2352b", #"Maximum Red"
           "#e58c31", # "Cadmium Orange"
           "#f7e336", #"Minion Yellow"
           "#fbf19b", # "Green Yellow Crayola"
           "#feffff", # "White"
           "#f1c2ae" # "Apricot"
)

```

## Movie Analysis

In this section we analyze the MCU as a whole - as it relates to budget, box office, ratings, and gender of main actors.

### Budget and Box Office

```{r, fig.width=12, fig.height=8}

avg_intl <- film_clean %>%
  filter(MCU == "Non-MCU", type == "Movie") %>%
  summarize(avg = mean(International, na.rm = TRUE)) %>%
  pull(avg)

avg_dom <- film_clean %>%
  filter(MCU == "Non-MCU", type == "Movie") %>%
  summarize(avg = mean(Domestic, na.rm = TRUE)) %>%
  pull(avg)

mcu_box_stack <- film_clean %>%
  filter(MCU_film == 1) %>%
  pivot_longer(cols = c(International, Domestic), names_to = "bo_loc", values_to = "boxoffice") %>%
  select(c(title, bo_loc, boxoffice)) %>% distinct() %>%
  ggplot(aes(x = reorder(title, boxoffice), y = boxoffice, fill = bo_loc)) + 
  geom_bar(position="stack", stat="identity") +
  scale_y_continuous(expand = c(0, 0), labels = label_dollar(suffix = " B", scale = 1e-9)) +
  geom_hline(yintercept = avg_intl, linetype = "dashed", size = 0.5, color = colors[7]) +
  #annotate("text", y = 2000000000, x = "Eternals", label = "Average International Gross", size = 3, color = colors[7]) +
  coord_flip() +
  scale_fill_manual(values = c(colors[11], colors[5]), name = "")+
  labs(title = "Box Office MCU Movies",
       subtitle = "Relative to Average International Box Office",
       y = "",
       x = "") +
  mcu_theme()

mcu_box_stack

```

Here we can see the breakdown of international and domestic box office numbers across MCU films. Endgame is far above the others, at almost $3 billion world wide gross, , followed by Infinity War and Spiderman: No Way Home. Ten of the movies gross above $1 billion. The orange dashed line represents international gross of the non-MCU movies these actors are in. All of the MCU films have higher international gross rates than that of the average non-MCU film, many tend to dwarf this figure. We can see just how much blockbusters, particularly the MCU, has changed the film industry and the definition of a successful film franchise. This is often a point of contention among film critics - how far should fandom go and is the Marvel machine never-ending? Are movies like those of the MCU pushing out smaller filmmakers, or is there room for both in cinema? 


```{r}
# Include just MCU movies

MCU_films <- film_clean %>%
  filter(MCU == "MCU") %>%
  filter(type == "Movie") %>%
  group_by(title) %>%
  summarize(BO = World_Wide, Budget = budget_USD, release_date = release_date) %>%
  distinct(title, BO, Budget, release_date) %>%
  mutate(Phase = as.factor(ifelse(release_date < as.Date("2012-05-05"), "One", ifelse(release_date < as.Date("2015-07-18"), "Two", ifelse(release_date < as.Date("2019-07-03"), "Three", "Four")))))

MCU_films$Phase <- relevel(MCU_films$Phase, "One")

```


```{r, fig.width=12, fig.height=8}
tidy_BOB <- MCU_films %>%
  dplyr::select(BO, Budget, Phase) %>%
  rownames_to_column("ID") %>%
  rename(Film_Budget = Budget, Box_Office = BO) %>%
  gather(var, values, -ID, -title, -Phase)


standardize <- function(x) (x-mean(x))/sd(x) 
df2 <- tidy_BOB %>% group_by(var) %>%   
  mutate(values = standardize(values)) %>% ungroup()


df2$var <- factor(df2$var, levels = c('Film_Budget', 'Box_Office'), labels = c('Film Budget', 'Box Office'))

gslope <- ggplot(df2, aes(x= var, y= values, group= ID, color= Phase)) +
  geom_line(size = 2) +
  scale_color_manual(values = c(colors[4], colors[2], colors[11], colors[6]))+
  labs(title= "Box Office Success vs Film Budget", 
       x= "", 
       y= "Z-Score Scaled") +
  mcu_theme()

gslope

```

Here we look at the relationship between the film's budget and its Box Office success across all of the MCU movies. When making a movie, the hope would typically be that if you invest more resources into making the film, you will reap the benefits of greater returns. For the most part, when it comes to the MCU films, that is generally true. However, it is definitely the case that there are other factors at play when it comes to Box Office success.
It is particularly interesting to look at the Phase 4 MCU films here. It appears as though each Phase 4 film had the same budget to work with, yet one movie received a disproportionately high Box Office return, while the other three had some of the lowest Box Office to Budget ratios. Clearly, budget was not the main differentiating factor here. One major determinant here was the Covid-19 pandemic. Stay-at-home and isolation orders caused a major blow to the cinema industry, as far fewer people were going out to see any movies. Thus, movies released during that time were destined to fare worse in the Box Office. Also during this time, possibly as a response, Disney made the decision to begin releasing their MCU movies on their streaming service, Disney+, just about 45 days after the theatrical release of the movie. This further incentivized potential movie-goers to stay home, because it wouldn't be long before they'd be able to see the movie at home anyway, so they wouldn't feel as though they were missing much by skipping the theater trip.
The final Phase 4 film, Spider-Man: No Way Home, conversely, had nearly the opposite situation. This movie's release came after the initial pandemic protocols had larger passed. Further, since the Spider-Man property is still in a bit of a tangle between Disney and Sony, No Way Home would not be coming to Disney+ any time in the near future. This, in turn, convinced many people who had been content to skip over the theater experience to feel as though this was a film they couldn't miss. Adding to that, the Spider-Man character was already one that generated more excitement than a few of the protagonists from the MCU's immediately preceding movies. All those factors combined to result in a Box Office explosion of nearly two billion dollars.
Also, it is worth noting that the majority of the MCU films with the highest budget are the big team up films, including the four Avengers movies, as well as Captain America: Civil War, which featured much of the Avengers team. So, Disney might know ahead of time that those particular movies will draw in a larger audience, and thus know that they need to increase the budget of those films to guarantee they are putting out a higher quality product.
Thus, while there might be a slight correlation between budget and Box Office success, there are also other factors at play that suggest that simply throwing a massive budget towards a film doesn't guarantee success at the Box Office.

```{r, fig.width=8, fig.height=6}
# all movies Box Office Success, Colored by MCU

gT <- ggplot(film_clean, aes(x = release_date, y = World_Wide, color = MCU, text = title)) +
  geom_point() +
  scale_color_manual(values = c(colors[2], colors[11])) +
  scale_y_continuous(labels = label_dollar(suffix = " B", scale = 1e-9)) +
  labs(title= "World-Wide Box Office Success", x= "Release Date", y= "World-Wide Box Office") +
 theme_minimal()

 ggplotly(gT)

```

Here we look at the Box Office success of the combined filmography of all the actors in our study. Before Iron Man kicked off the MCU in 2008, we can see that a few of these actors appeared in a handful of blockbuster movies, including entries from the Jurassic World and Star Wars universes. However, once the MCU established themselves in the movie landscape, we see that many of the blockbusters that our actors appeared in did in fact from the MCU. A few movies stand out far above the crowd, including Avengers: Endgame, which challenged Avatar for the highest Box Office total of any movie.
One of the many impressive feats of the MCU is their general lack of complete misfires. Even the lowest Box Office returns among the MCU movies still rank highly when compared to the majority of films that these actors appear in. Suffice it to say, when an actor signs on for a role in an MCU movie, they can be quite certain that the film will receive a much higher Box Office return than the average non-MCU film.

### Movie Ratings

```{r, fig.width=12, fig.height=6}

hex_ratings <- ggplot(film_clean, aes(x = averageRating, y = World_Wide)) +
  geom_hex() +
  labs(x = "Average Rating",
       y = "Box Office (world-wide)",
       title = "Ratings and Budgets are higher for MCU films") +
  scale_fill_gradient(low = colors[4], high = colors[2], name = "Movie Count") +
  scale_y_continuous(labels = label_dollar(suffix = " B", scale = 1e-9)) +
  facet_wrap(~MCU)+
  mcu_theme() +
  theme(strip.text.x = element_text(size = 18))


hex_ratings

```

Here we can see the distribution of ratings and box office for MCU and non-MCU movies (all from actors in the MCU). We can see that ratings for MCU movies are all above 6. For non-MCU movies, although most of the ratings are above 6, there are quite a few under that. We can also see a cluster of darker blue in the non-MCU chart with low box-office numbers. Although again, the scale is somewhat distorted due to the very high grossing films in the MCU and Avatar, not in the MCU. 


### Gender of Main Actors & Awards

```{r, fig.width=12, fig.height=8}

gender <- film_clean %>%
  filter(MCU == "Non-MCU", type == "Movie") %>%
  mutate(Awards = ifelse(is.na(winner), "None", winner),
         Awards = ifelse(Awards == "Nominated", "Nom", Awards),
         Awards = ifelse(Awards == "Winner", "Won", Awards)) %>%
  group_by(actor_name, Gender, Awards, `BeforeMCU`) %>%
  summarize(Freq = n(), .groups = 'drop') %>%
  ungroup() %>%
  group_by(Gender, `BeforeMCU`, Awards) %>%
  summarize(Freq = mean(Freq), .groups = 'drop') %>%
  ungroup() %>%
  mutate(Freq = ceiling(Freq),
         BeforeMCU = factor(BeforeMCU)) 

# mosaic plot

mosaic(BeforeMCU ~ Gender + Awards, direction = c("v", "v", "h"), gender, 
       highlighting_fill = c(colors[4], colors[2]),
       rot_labels = c(0, 0, 90, 0),
       gp_varnames = gpar(fontsize = 22, fontface = 1),
       gp_labels = gpar(fontsize = 18))

```

From this gender breakdown, we can first see that there are more men in the MCU than women. We can also see that the average number of films actresses have done before entering the MCU exceeds that of men, but not by much. For both men and women, all Oscars were won before doing an MCU film. Women in the MCU have slightly more award nominations ("Nom") than men after being in the MCU. The MCU mainly highers well-known actors, so this breakdown may not be surprising. Although it could suggest that actors are less likely to win an Oscar after participating in an MCU film. More likely it is that MCU films take up large amounts of time for actors, and while the payout exceeds that of other types of films, actors are not likely to be nominated or win Oscars for blockbuster movies like these, despite excellent performances. We do see some Emmy nominations for the more recent MCU TV shows - Wandavision and Falcon and the Winter Soldier were both nominated for a few awards, but did not win any. 


## Actor Analysis

In this section we analyze the actors in the MCU - as it relates to their success across films within and outside of the MCU.

### Actor Filmography

```{r, fig.width=14, fig.height=8}

g1 <- film_clean %>%
  filter(year >= 1990) %>%
  filter(type == "Movie") %>%
  ggplot(aes(x = release_date, y = actor_name, color = MCU, size = World_Wide)) +
  scale_size(range = c(3, 8), name = "Box Office", breaks = c(1000000, 500000000, 1000000000, 2000000000),
             labels = c("$1M", "$500M", "$1B", "$2B")) +
  scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  scale_color_manual(values = c(colors[2], colors[11]))+
  labs(y = "Actor Name",
       x = "Release Date",
       title = "Filmography Box Office")+
  geom_point(alpha = 0.8, stroke = 0) +
  mcu_theme() +
  guides(colour = guide_legend(override.aes = list(size=10)))

g1

```

For most of the actors, the MCU films are the highest grossing films that they are a part of. For some actors, there are a few other notably high grossing films, such as Star Wars, Jurassic World, Avatar, and the Hobbit movies. MCU box office numbers tend to dwarf those of other films, so the comparison is somewhat difficult. There is not a clear pattern indicating that actors who work on MCU movies, then follow with smaller scale productions. Although it does seem that after Endgame (the last very large blue dot form 2019), many of the actors have participated in far fewer films than they did before. This could also be due to Covid slowing the release dates of many productions.


```{r, fig.width=14, fig.height=10}

awards_graph <- film_clean %>%
  filter(year >= 1990) %>%
  filter(type == "Movie") %>%
  arrange(award_mcu, descending = FALSE)
  
  
g_awards <- ggplot(awards_graph) +
  geom_point(mapping = aes(x = release_date, y = actor_name, color = award_mcu, text = title), 
             size = 4, alpha = 0.8) +
  scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  scale_color_manual(values = c(colors[4], colors[2], colors[11], colors[6]), name = "Film Type")+
  labs(y="",
       x = "Release Date",
       title = "Film Releases by Actor",
       subtitle = "Any Credited Role")+
  mcu_theme() +
  guides(colour = guide_legend(override.aes = list(size=10)))

g_awards

```

Most actors have quite a few roles before entering the MCU (only showing after 1990). The MCU tends to bring on more known actors - every actor that has won an Oscar has done so before being in an MCU film. There are a few actors Zendaya, Tom Hiddleston, Chris Hemsworth, who had sparse movie credits before entering the MCU. Zendaya is more famous for her role as Rue in Euphoria, for which she was the youngest actress to recieve the Emmy for best actress. 


```{r, fig.width=12, fig.height=8}

h1 <- film_clean %>%
  filter(type == "Movie") %>%
  group_by(actor_name, decade) %>%
  summarize(movie_count = n(), .groups = 'drop') %>%
  arrange(decade) %>%
  ggplot(aes(x = decade, y = actor_name, fill = movie_count)) +
  geom_tile(color = "grey") +
  #scale_y_discrete(limits=rev) +
  coord_fixed(ratio = 0.5) +
  scale_fill_gradient(low = colors[4], high = colors[2], name = "Movie Count") +
  labs(y = "Actor",
       x = "Decade",
       title = "Movie Count by Decade") +
  mcu_theme() +
  theme(panel.grid = element_blank())
  
h1

```

Here we look at the count of movies for each actor across the decades. Michael Douglas is the only actor with roles dating back to the 60's. Samuel L. Jackson appeared in many movies over the course of his career. Most actors appear in the most number of movies in the 2000's and 2010's, which makes sense since Marvel movies started in 2008. Most actors did not start acting in movies until the 2000's.   


```{r, fig.width=12, fig.height=8}
first_credit <- film_clean  %>%
  filter(!title %in% c("The Simpsons", "The Tonight Show with Jay Leno", 
                       "Late Night with Conan O'Brien", "Sesame Street")) %>%
  group_by(actor_name) %>%
  slice(which.min(release_date))

g_first <- first_credit %>%
  ggplot(aes(x = release_date, y = reorder(actor_name, release_date), color = type)) +
  geom_segment(aes(x = min(release_date), xend = as.Date(release_date), y = reorder(actor_name, release_date), 
                   yend = reorder(actor_name, release_date)), color = "grey") +
  geom_point(size = 5) +
  scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  scale_color_manual(values = c(colors[6], colors[11]), name = "")+
  labs(title = "Most MCU actors got their start in TV",
       subtitle = "First Credited Role",
       y="",
       x = "Release Date")+
  mcu_theme()

g_first
```

For most actors in the MCU, their first credited role is on a TV show. This makes sense for actors starting young or having a small role on a show. Movies are somewhat clustered for release dates, there is a cluster of first appearances in movies around the late 80's - 90's and the three most recent debuts were in movies. 


```{r}
# Add First MCU Film Date for Each Actor
actor_change1 <- film_clean %>%
  filter(type == "Movie") %>%
  group_by(actor_name, MCU) %>%
  arrange(actor_name, release_date) %>%
  filter(release_date == min(release_date)) %>%
  mutate(first_MCU = release_date) %>%
  filter(MCU == "MCU") %>%
  ungroup() %>%
  dplyr::select(c(actor_name, first_MCU))


First_MCU_added <- film_clean %>%
  left_join(actor_change1) %>%
  mutate(joining_phase = ifelse(first_MCU < as.Date("2012-05-05"), "_One", ifelse(first_MCU < as.Date("2015-07-18"), "Two", ifelse(first_MCU < as.Date("2019-07-03"), "Three", "Four")))) %>%
  mutate(joining_phase = factor(joining_phase, levels= c("_One", "Two", "Three", "Four")))

First_MCU_added$joining_phase <- relevel(First_MCU_added$joining_phase, "_One")

```

### Box Office Success & Ratings

```{r, fig.width=12, fig.height=8}
# Actor Success Before and After First MCU Film

avg_success <- film_clean %>%
  filter(MCU == "Non-MCU", type == "Movie") %>%
  group_by(actor_name, BeforeMCU) %>%
  summarize(avg_bo = mean(World_Wide, na.rm = TRUE)) %>%
  ungroup() %>%
  pivot_wider(names_from = BeforeMCU, values_from = avg_bo)

gBaA <- ggplot() +
  geom_segment(avg_success, mapping = aes(x = reorder(actor_name, Before), xend = actor_name, y = Before, yend = After), color = "grey") +
  geom_point(avg_success, mapping = aes(x = reorder(actor_name, Before), y = Before), size = 4, color = colors[5]) +
  geom_point(avg_success, mapping = aes(x = actor_name, y = After), size = 4, color = colors[7]) +
  annotate("text", y = 350000000, x = "Brie Larson", label = "Before MCU", size = 5, family = "Helvetica", hjust = 0) +
  annotate("text", y = 350000000, x = "Jeremy Renner", label = "After MCU", size = 5, family = "Helvetica", hjust = 0) +
  geom_point(avg_success, mapping = aes(y = 340000000, x = "Brie Larson"), size = 8, color = colors[5]) +
  geom_point(avg_success, mapping = aes(y = 340000000, x = "Jeremy Renner"), size = 8, color = colors[7]) +
  coord_flip() +
  scale_y_continuous(labels = label_dollar(suffix = " M", scale = 1e-6)) +
  scale_color_manual(name = "Before/After",
                     values = c("Before" = colors[5], "After" = colors[7]),
                     breaks = c("Before", "After")) +
  labs(title= "Average Box Office Success Before vs After First MCU Film", 
       x= "Actor Name", 
       y= "Average Box Office Success (non-MCU)") +
  mcu_theme()

gBaA

```

Here we look at the Average Box Office success of each actor before and after their first MCU film. This graph does not include the values from any MCU movies themselves. For example, even though Iron Man 2 came out after Iron Man, and would thus count towards Robert Downey Jr's average Box Office success after his first MCU film, we did not include it in calculating the average. That choice was made because the Box Office numbers for pretty much all MCU movies are through the roof and would skew the data to heavily favor the "after" average. The idea was to see if the movies these actors were in were generally more successful after these actors made their appearance in the MCU, so for that reason we wanted to separate them from the MCU films themselves that are destined to do well in the Box Office.
With that distinction made, we turn to the data. It turns out, there does not seem to be much of a trend in either direction. There are a few extreme cases in both directions, but ultimately it does not appear that there is any guarantee that appearing in an MCU film will make people all flock to the next non-MCU film starring that actor. Possibly the biggest increase that we see is from Chris Pratt, who first appeared as Star-Lord in Guardians of the Galaxy, back in 2014. The average Box Office return for movies featuring Chris Pratt approximately quadrupled after he made his first MCU appearance. However, we once again want to look a bit deeper into the underlying reasons as to why that is. In looking at Pratt's filmography, we see that he starred in a few of the Jurassic World movies after his MCU debut. The Jurassic World universe was already an established film series, and was set to be a blockbuster that brought in impressive Box Office numbers. Was that success increased as a result of Chris Pratt's fame from starring in an MCU movie, or were the numbers for those movies always going to be high? In truth, there is probably a little bit of both.
On the reverse end of the spectrum, Evangeline Lilly saw the biggest decrease in average Box Office success. Her case is nearly the opposite of Chris Pratt's. Having appeared in the Hobbit movies before her first MCU film, her average beforehand was actually the highest of any actor in this study. Additionally, she has only appeared in a couple smaller films after her MCU debut in Ant-Man. As such, she was our biggest faller.
Overall, though, there is almost no trend to predict whether an actor will increase or decrease in their average Box Office returns after debuting in an MCU film.



```{r}
# Parallel Coordinate Plot for Actors and Various Stats

award_nom_count <- film_clean %>%
  group_by(actor_name) %>%
  summarize(Award_Nominations = length(na.omit(winner)))


master_MCU <- film_clean %>%
  filter(MCU == "MCU")

master_Non <- film_clean %>%
  filter(MCU == "Non-MCU")

MCU_film_total <- master_MCU %>%
  group_by(actor_name) %>%
  summarize(MCU_film_count = length(title), MCU_Avg_BoxOffice = mean(World_Wide, na.rm= TRUE), MCU_Avg_Budget = mean(budget_USD, na.rm= TRUE), MCU_Avg_Rating= mean(averageRating, na.rm= TRUE))

Non_film_total <- master_Non %>%
  group_by(actor_name) %>%
  summarize(film_count = length(title), Avg_BoxOffice = mean(World_Wide, na.rm= TRUE), Avg_Budget = mean(budget_USD, na.rm= TRUE), Avg_Rating= mean(averageRating, na.rm= TRUE))

actor_joining_phase <- First_MCU_added %>%
  group_by(actor_name) %>%
  summarize(joining_phase = joining_phase)


PC_data <- award_nom_count %>%
  left_join(MCU_film_total) %>%
  left_join(Non_film_total) %>%
  left_join(actor_joining_phase)

```




```{r}
#ggparcoord(PC_data, columns= c(2, 3, 7, 8, 4, 5, 9, 10, 6), alphaLines= .1, scale= "uniminmax", groupColumn= 11)  +
  #labs(title= "Actor Comparison Over Multiple Variables") +
  #theme_minimal() +
  #theme(title = element_text(family = "Helvetica", size = 20),
   #     axis.title = element_text(family = "Helvetica", size = 18),
    #    axis.text.x = element_text(angle= 60, family = "Helvetica", size = 10),
     #   axis.text.y = element_text(family = "Helvetica", size = 16),
      #  legend.title = element_text(family = "Helvetica", size = 18),
       # legend.text = element_text(family = "Helvetica", size = 16)
        #)

```


```{r, fig.width=20, fig.height=4}
PC_data$joining_phase <- relevel(PC_data$joining_phase, "_One")

PC2_data <- PC_data %>%
  dplyr::select(c(actor_name, joining_phase, MCU_film_count, film_count, Avg_BoxOffice, Avg_Budget, Avg_Rating, MCU_Avg_BoxOffice, MCU_Avg_Budget,  MCU_Avg_Rating, Award_Nominations))

colnames(PC2_data) <- c("Actor", "Joining Phase", "MCU Films", "Total Films", "Avg Box Office", "Avg Budget", 
                       "Avg Rating", "Avg MCU Box Office", "Avg MCU Budget",  "MCU Avg Rating", "Award Nominations")

PC2_data %>%
  parcoords(rownames= FALSE, brushMode = "1D-axes", reorderable = TRUE, queue = TRUE, alpha= .5, 
            color= list(colorBy = "Joining Phase", colorScale= "scaleOrdinal", colorScheme= c(colors[4], colors[2], colors[11], colors[6])), 
            withD3 = TRUE, width = 1100, height= 400
            )
```


Here we play around with quite a few relationships for each actor. First, it's worth noting that the lines are colored by when the actor made their first MCU appearance, a different color indicating each of the four "Phases" in the MCU. Of course, it isn't much surprise that the actors who first joined in Phase 1 generally have the most MCU film appearances, as they've had the most opportunities to be in them. This trend sort of continues into total film appearances (not including MCU films). Of the actors who got their start in Phase 3 or 4, they are mostly near the bottom of the total films list. This indicates that for the new actors that the MCU has been bringing in recently, they have been generally hiring lesser known actors compared to those who have already had an extensive film career. Speaking of extensive film careers, it was fun to note just how many films Samuel L Jackson has appeared in. He is so far ahead in the total film count that he looks to be an outlier in the data.
There actually appears to be a little bit of clustering among the actors who joined in Phase 1. This could be the result of sharing many of their MCU movies together. A lot of those actors ended up appearing in all of the Avengers movies, and so the MCU related stats ended up being very similar, the only differences coming from any difference in the actors' respective standalone films, but even those don't have quite as much effect on the data here because of all the various Avengers films adding weight to the average.
There does not appear to be too much correlation between the average Box Office success for actors from their MCU movies to their non-MCU movies. Thus it doesn't appear as though featuring in the highest performing MCU movies will mean that the non-MCU movies will also be as successful, and vice versa. Similarly, there does not seem to be any major trend between the average ratings for each actor's MCU films to their non-MCU films.



